sensi_an_rotations1. Start of sensitivity analysis for rotations.

### 2018-05-08 inst/documents/paper3/paper3_rotations.Rmd
paper3 writing

```{r, eval=TRUE, include=FALSE}
#!!!! PAPER PROTOCOL
#!!!! submission MS, knit to word_document 
#!!!!  in WORD: delete figs, set double space, add line numbers 
#!!!!  (spacing & line numbers work for pdf output but not for word)
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='pap2figs/')
```

@@

A population genetic model was developed simulating changes over time in the frequencies of resistant and susceptible alleles in response to fitness differences. The model uses a similar approach that has been used to compare insecticide mixtures and sequences [@Levick2017,@South2018]. The implementation is simpler here because for rotations and sequences there is only a need to follow one insecticide at a time. The simpler implementation allows the model to be run for an unlimited number of insecticides. The algebra behind the model is described in the supplementary information.

@@

2. Rotations : use insecticide for set time, switch to another.

In both cases the choice of the new insecticide to use was made by going through the list of available insecticides and choosing the next one for which resistance was below the threshold. Thus there is the potential to switch back to insecticides that had been used earlier in a model run if resistance had not reached the threshold or if resistance had declined below the threshold during a period when not in use.

In both cases a simulation run continued until no more insecticides remained below their resistance thresholds.

@@

@@

1. no resistance fitness costs or dispersal link to untreated areas
2. fitness costs of resistance added
3. dispersal link to untreated areas (no costs)
4. costs and dispersal

To generate a decline in resistance over time when an insecticide is not being used either fitness costs or a dispersal link to an untreated area are required.

@@

It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative. For this we summed the number of generations when an insecticide was deployed and the resistance allele frequency for that insecticide was below 0.5. This allowed comparison between strategies using a variable number of insecticides. Previously time-to-resistance, namely the number of generations it takes to reach a defined resistance threshold (usually a resistance allele frequency of 0.5), was used. Using time-to-resistance worked when considering just two insecticides, but when the number of insecticides is a variable it is not sufficient.

@@

Example model runs are shown with no costs or dispersal (Fig 1), with just resistanace costs (Fig 2) and with just dispersal from untreated refugis (Fig 3).

#### Results1. no resistance fitness costs or dispersal from untreated refugia

When there are no resistance fitness costs or dispersal links to untreated refugia resistance frequencies do not decline when an insecticide is not in use (Fig 1). Thus for sequences once resistance thresholds are reached the resistance frequencies remain at that level and the insecticide cannot be re-used. For rotations resistance frequencies step upwards when insecticides are in use and remain at a plateau when they are not.

@@

#### Results2. fitness costs of resistance added

With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold.

In contrast to the previous section 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).

#### Results3. dispersal from untreated refugia added

@@

#### Results4. costs and dispersal

@@

#### Runs where rotations have substantial advantage over sequences
As shown in Fig 4.2 some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 5.

@@

@@

@@

@@

The model does not (although it could) include the potential effects of modifier genes ameliorating the costs of resistance alleles. It has been suggested that rotations could be favoured as a strategy because they keep resistance frequences lower and restrict selection pressure for modifier genes [@WHO2012]. However, there is little evidence to support the existence of such modifier genes in insectide resistance []. Even with the existence of modifier genes it is unlikely that they would provide much of an advantage to rotations in the scenarios we describe. In the modelled scenarios the use of an insecticide is stopped when the resistance frequency for that insecticide reaches 0.5. Thus in all scenarios resistance frequencies do not remain at high levels for long. In this situation it seems unlikely that modifiers would be selected for.

@@

## TODO
*try to come up with brief name for model output to make it easier to refer to it later*

@@

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence are stopped at the same time. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and remains at the constant level while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence."}

@@

```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4. Effect of cost and dispersal on the distribution of differences between rotations and sequences from model runs. The y axis is square root transformed to emphasise the infrequent runs where there is a significant benefit of rotations over sequences. 4.1 when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results. 4.2, 4.4 when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences. For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences. 4.3 when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}

@@

  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=pcent_rot_adv)) +

             geom_histogram(binwidth=10) +
             xlim(-20,400) +
             xlab('') +
             ylab('count') +
             ggtitle('1. no resistance costs or dispersal') +
             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
             theme_minimal() +
             theme(panel.grid.minor.y=element_blank())

  gg_cost <- ggplot( df_cost, aes(x=pcent_rot_adv)) +

             geom_histogram(binwidth=10) +
             xlim(-20,400) +
             xlab('') +
             ylab('count') +
             ggtitle('2. with costs') +
             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
             theme_minimal() +
             theme(panel.grid.minor.y=element_blank())

  gg_mig <- ggplot( df_mig, aes(x=pcent_rot_adv)) +

             geom_histogram(binwidth=10) +
             xlim(-20,400) +
             xlab('') +
             ylab('count') +
             ggtitle('3. with dispersal refugia') +
             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
             theme_minimal() +
             theme(panel.grid.minor.y=element_blank())

  #gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_minus_seq)) + 
  gg_cost_mig <- ggplot( df_cost_mig, aes(x=pcent_rot_adv)) +              
             geom_histogram(binwidth=10) +
             xlim(-20,400) +
             xlab('percent benefit of rotation versus sequence') +
             ylab('count') +
             ggtitle('4. with costs and dispersal') +
             scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + #can't use log scale because anchored at 0
             theme_minimal() +
             theme(panel.grid.minor.y=element_blank())

@@

```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 6. Which model inputs have the greatest influence on the difference between rotations and sequences when costs alone are included ? Measured using Partial Rank Correlation Coefficients (PRCC). Dominance of cost and dominance of selection have the greatest influence."}

@@

```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 7. Model runs with substantial benefits of rotations over sequences are associated with high dominance of cost and low dominance of selection. Dominace inputs are shown on the x and y axes. Small grey points indicate all model runs. Model runs in which rotations last greater than 20% longer than sequences are shown in colour with darker blue indicating those that have the greatest difference."}

>>>>

```{r, eval=TRUE, include=FALSE}
#!!!! PAPER PROTOCOL
#!!!! submission MS, knit to word_document 
#!!!!  in WORD: delete figs, set double space, add line numbers 
#!!!!  (spacing & line numbers work for pdf output but not for word)
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='pap3figs/')
```

@@

A population genetic model was developed tracking the change in resistance allele frequency as consequence of the genetic determination of the resistance phenotype, and the deployment policy for the insecticides. The model uses a similar approach to that previously used to compare insecticide mixtures and sequences [@Levick2017,@South2018]. The implementation is simpler here because for rotations and sequences there is only a need to follow one insecticide at a time and linkage disequilibrium can be ignored. The simpler implementation allows the model to be run for an unlimited number of insecticides. The algebra behind the model is described in the supplementary information.

@@

2. Rotations : use insecticide for set time (i.e. number of insect generations), switch to another.

In both cases the choice of the new insecticide to use was made by going through the list of available insecticides and choosing the next one for which resistance was below the threshold. Thus there is the potential to switch back to insecticides that had been used earlier in a model run if resistance had not reached the threshold or if resistance had declined below the threshold during a period when not in use. This means that the “sequential” policy is effectively a “rotate when resistant” policy but we use the former term for consistency with previous work. *todo cite different trems from Rex consortium review* For both rotations and sequences the simulation continues until no more insecticides remain below their resistance thresholds.

@@

**Table 1. Model inputs and the ranges used in simulations**

Input  | Description | Min | Max
---------------------- | ------------------------------------- | ----- | -----
Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide | 0.5 | 1
Exposure | proportion of females exposed to insecticide | 0.4 | 0.9
Male exposure | proportion of males exposed to insecticide as proportion of females | 0 | 1
Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide | 0.1 | 0.9
Dominance of resistance | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide | 
Frequency  | frequency of resistance alleles within the population | 0.005 | 0.1
Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide | 0 | 0.1
Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide | 0.1 | 0.9
Insecticide number | number of independent insecticides available | 2 | 5  
Rotation interval | number of generations to use insecticide before switching | 5 | 50
Coverage | for scenarios with dispersal, the proportion of insects in the treated area rather than untreated refugia |  0.1 | 0.9
Dispersal | proportion of population exchanged between treated and untreated areas per generation, 0=none, 1=random mixing |  0.1 | 0.9

@@

1. baseline : no resistance fitness costs or dispersal to/from untreated refugia
2. fitness costs of resistance added
3. dispersal link to untreated areas (no costs)
4. costs and dispersal

For each scenario 10,000 runs were performed randomly selecting inputs according to the ranges specified in Table 1 using a uniform distribution in all cases.

@@

*todo change this bit*
It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative. For this we summed the number of generations when an insecticide was deployed and the resistance allele frequency for that insecticide was below 0.5. This allowed comparison between strategies using a variable number of insecticides. Previously time-to-resistance, namely the number of generations it takes to reach a defined resistance threshold (usually a resistance allele frequency of 0.5), was used. Using time-to-resistance worked when considering just two insecticides, but when the number of insecticides is a variable it is not sufficient.

@@

Example model runs are shown with no costs or dispersal (Fig 1), with just resistance costs (Fig 2) and with just dispersal from untreated refugia (Fig 3).

*todo describe these example runs first, before talking about overall results*

#### Results1. form of resistance evolution with and without costs and dispersal.

When there are no resistance fitness costs or dispersal links to untreated refugia resistance frequencies do not decline when an insecticide is not in use (Fig 1). Thus for sequences once resistance thresholds are reached the resistance frequencies remain at that level and the insecticide cannot be re-used. For rotations resistance frequencies step upwards when insecticides are in use and remain at a plateau when they are not. If the rotation interval is short enough then the insecticide can be used a few times before the resistance thresholds are reached.

### Results2 : comparing rotations and sequences across model runs.

In all the example runs shown the resistance threshold (allele frequency of 0.5) was reached within the course 500 generations which was the maximum limit imposed in the model. Across all model runs there were many in which the thresholds were not reached in this time period.

*but first need to explain that sequences pretty much never better*
*so show the modified histograms first, i.e. don't filter out the threshold not reached runs*

Thus to aid the comparison of rotations and sequences model runs can be broadly classified into three groups :

@@

*todo : maybe I should just leave the 'succeeded' runs in here which would change these %s slightly*
With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold. In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).

@@

@@

#### Runs where rotations have substantial advantage over sequences

*Talk about classing runs into both 'succeeded', both 'failed' or just rotation succeeded.
*show violin plots here ?*

As shown in Fig 4.2 some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 5.

@@

This can be understood intuitively. When there are no costs, the dynamics are the same, it is simply that sequential deployment proceeds in a few, larger steps while rotations proceeds in more numerous smaller steps; both reach the same point in around the same time (e.g. figs 1 and 3). However dominance effects on cost and resistance can act to keep frequencies low in a rotations policy. Recall that at low frequencies of resistance most resistance alleles are present as heterozygotes RS and only a vey small proportions heterozygotes. However costs are dominant and still act at low frequencies. This is illustrated on Figure Y. Relatively rapid rotations keep the allele frequencies at the LH side of Figure Y where costs during the periods when the insecticide is not deployed greatly outweigh the selection for IR during the periods when the insecticide is deployed. We believe the explanation is therefore that rotations keep allele frequencies in the parameter space to the left of Figure Y and hence acts to slow the evolution of IR compared to sequences which allow frequencies to rise to relatively high levels before the insecticide is replaced.  (does this sound reasonable??). Note also that starting frequency, as expected, had an impact (Figure 7), lower starting frequencies slightly favouring rotations by allowing more time in the “low frequency” parameter space.

@@

Note however that this combination of recessive resistance and high dominance costs are exactly the features most desirable in any insecticide irrespective of how it is deployed. We discuss fitness below but note here that it is very difficult to maintain insecticide concentration at the level required to make resistance recessive. Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant; see Figure 1 of <cite Levick et al> and our more extensive discussion in <cite South et al WoS manuscript).

@@

We also found frequency of rotations had little effect. This is important in the control of vectors of human diseases such as malaria and dengue. These predominantly occur in resource-poor areas often with poor healthcare infrastructure. Experience suggests that rotation targets would often be missed (e.g rotations occur approximately every 15 generations rather than the planned “official” 10 generations) but our simulations suggest this would not fatally undermine a rotations policy.

@@

The model does not include the potential effects of modifier genes ameliorating the costs of resistance alleles. To include this would require a more detailed model tracking linkage disequilibrium between each resistance locus and its modifier. It has been suggested that rotations could be favoured as a strategy because they keep resistance frequences lower and restrict selection pressure for modifier genes [@WHO2012]. However, there is little evidence to support the existence of such modifier genes in insectide resistance []. Even with the existence of modifier genes it is unlikely that they would provide much of an advantage to rotations in the scenarios we describe. In the modelled scenarios the use of an insecticide is stopped when the resistance frequency for that insecticide reaches 0.5. Thus in all scenarios resistance frequencies do not remain at high levels for long. In this situation it seems unlikely that modifiers would be selected for.

@@

In summary, we show that rotations are invariably better than sequential use, although the difference may often be small (figure 4); certainly rotations should not be seen as a panacea capable of removing the threat of resistance evolution. Rotations may have a large impact if fitness costs are prints and dominant (these are genetic/physiological factors outside our control)  and if resistance can be kept recessive. The latter is under our control to some extent (fig 1 of Levick et al ) but it is extremely difficult to deploy and maintain insecticide concentrations at levels that ensure recessively (cite WoS ms): as concentrations decline resistance become dominant and any befits of rotations largely disappear.

## TODO
*try to come up with brief name for model output to make it easier to refer to it later*
*add caveat about polygenic resistance*
*The caption to each graph needs to have the list of parameter values used to generate them*
*work on Ians fig Y to help explanation*

@@

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia. The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence reach the endpoint, resistance frequency > 0.5 to all insecicides, at the same time. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and remains at the constant level while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. All other inputs are kept constant between the rotation and the sequence."}

@@

```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4. Effect of cost and dispersal on the difference between rotations and sequences from model runs. The y axis shows the number of model runs for the different caegories of difference between rotations and sequences. 4.1 when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results. 4.2, 4.4 when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences. For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences. 4.3 when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}

@@

  # log scale overemphasised low values
  # sqrt scale looked good but Ian didn't like

  custom_bins <- c(-20,-1,1,20,100,450)
  df_nocost_nomig$rot_adv_bins <- cut(df_nocost_nomig$pcent_rot_adv, custom_bins)
  df_cost$rot_adv_bins <- cut(df_cost$pcent_rot_adv, custom_bins)
  df_mig$rot_adv_bins <- cut(df_mig$pcent_rot_adv, custom_bins)
  df_cost_mig$rot_adv_bins <- cut(df_cost_mig$pcent_rot_adv, custom_bins)

  bin_labs <- levels(df_cost$rot_adv_bins)

  #gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=pcent_rot_adv)) + 
  gg_nocost_nomig <- ggplot( df_nocost_nomig, aes(x=rot_adv_bins)) +

             geom_bar() +
             #xlim(-20,400) +
             xlim(bin_labs) +
             xlab('') +
             ylab('count') +
             ggtitle('1. no resistance costs or dispersal') +
             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
             theme_minimal() +
             theme(panel.grid.minor.y=element_blank(),
                   axis.text.x=element_blank())

  gg_cost <- ggplot( df_cost, aes(x=rot_adv_bins)) +

             geom_bar() +
             #geom_histogram(binwidth=10) +
             #xlim(-20,400) +
             xlab('') +
             ylab('count') +
             ggtitle('2. with costs') +
             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
             #scale_x_continuous(breaks=c(-10,-1,0,1,20,100,400)) +
             scale_x_discrete( limits=bin_labs,
                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
             theme_minimal() +
             theme(panel.grid.minor.y=element_blank(),
                   axis.text.x=element_blank())

  gg_mig <- ggplot( df_mig, aes(x=rot_adv_bins)) +

             geom_bar() +
             xlim(bin_labs) +
             xlab('') +
             ylab('count') +
             ggtitle('3. with dispersal refugia') +
             theme_minimal() + 
             theme(panel.grid.minor.y=element_blank(),
                   axis.text.x=element_blank())

  #gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_minus_seq)) + 
  gg_cost_mig <- ggplot( df_cost_mig, aes(x=rot_adv_bins)) +

             geom_bar() +
             xlab('Percent benefit of rotation over sequence') +
             ylab('count') +
             ggtitle('4. with costs and dispersal') +
             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
             theme_minimal() +
             scale_x_discrete( limits=bin_labs,
                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+       
             theme(panel.grid.minor.y=element_blank())

@@

```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 6. Which model inputs have the greatest influence on the difference between rotations and sequences ? In this case for scenario 2 where resistance costs are included but dispersal is not. Measured using Partial Rank Correlation Coefficients (PRCC). Positive values indicate a positive influence on the benefit of rotations, only cost and dominance of cost in this case. Dominance of cost and dominance of selection have the greatest magnitude of influence."}

@@

```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 7. Model runs in which rotations perform more than 20% better than sequences are associated with high dominance of cost and low dominance of selection. Dominance inputs are shown on the x and y axes. Small grey points indicate all model runs. Model runs in which rotations last greater than 20% longer than sequences are shown in colour with darker blue indicating those that have the greatest difference."}

### 2018-05-08 inst/documents/rotations_notes.Rmd
paper3 writing

@@

~ run 10k runs of dispersal overnight
~ try histograms on log scale ?
~ try decision tree to predict rot_minus_seq from inputs

~ send MS to Ian before 11

>>>>

*2/5/18*

exchanging emails with Ian

Thanks Ian, I'm getting closer to understanding. I'm just trying to understand why the lines cross, which suggests to me that resistance wouldn't be selected for at low frequencies ?

This is a simple text description of how I'm starting to understand. Is it true ? (we can work on improving the wording later).

Resistance increases when an insecticide is in use and, if there are resistance costs, decreases when it is not.

If the difference between this increase and decrease over periods of use and non-use is positive then resistance will keep increasing.

If the positive difference between the increase and decrease is lower at lower frequencies then there is an advantage to short rotation intervals that keep frequencies low.

In this latter scenario sequential use leads to higher frequencies where increases during periods of use are more likely to outweigh decreases during periods of non-use.

done ~ Can I change histograms of count of rot advantage to have different bins, we're only interested in whether difference is 0, moderate or substantial.

Trying to recreate scenario 2167
after removal of non threshold runs :

  scenario <- 2167
  linputs <- as.list( df_cost[scenario, 1:15] )
  dfres <- do.call(run_rot, linputs)
  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
  linputs$rot_interval <- 0
  dfres <- do.call(run_rot, linputs)
  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE) 
  grid.arrange(ggrot, ggseq, ncol=1)

linputs$dom_cos <- 0.8
x11()

  linputs$rot_interval <- 10
  dfres <- do.call(run_rot, linputs)
  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
  linputs$rot_interval <- 0
  dfres <- do.call(run_rot, linputs)
  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE) 
  grid.arrange(ggrot, ggseq, ncol=1)

# setting params for 2167
linputs <- list()
linputs$n_insecticides <- 3
linputs$coverage <- 1
linputs$migration <- 0
linputs$cost <-     0.0582 
linputs$max_gen <-  500
linputs$expo_hi <-  0.64
linputs$eff <-      0.96
linputs$rr <-      0.78
linputs$dom_sel <-  0.06
linputs$dom_cos <-  0.7
linputs$start_freqs <- 0.0731
linputs$male_expo_prop <- 0.167
linputs$plot <- FALSE

  linputs$rot_interval <- 10
  dfres <- do.call(run_rot, linputs)
  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
  linputs$rot_interval <- 0
  dfres <- do.call(run_rot, linputs)
  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE) 
  grid.arrange(ggrot, ggseq, ncol=1)

# this shows a small tweek of the params leads to the sequence lasting forever too
# and I don't know which input change causes the seq to last forever ?
# Aha seems that I hadn't included rr=0.78, rr default is 0.5, 
# with lower rr less selection for r and sequence does better 
linputs <- list()
linputs$n_insecticides <- 3
linputs$coverage <- 1
linputs$migration <- 0
linputs$cost <-     0.058 
linputs$max_gen <-  500
linputs$expo_hi <-  0.64
linputs$eff <-      0.96
linputs$rr <-      0.5 #0.78
linputs$dom_sel <-  0.06
linputs$dom_cos <-  0.7
linputs$male_expo_prop <- 0.167
linputs$start_freqs <- 0.0731
linputs$plot <- FALSE

  linputs$rot_interval <- 10
  dfres <- do.call(run_rot, linputs)
  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
  linputs$rot_interval <- 0
  dfres <- do.call(run_rot, linputs)
  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE) 
  grid.arrange(ggrot, ggseq, ncol=1)

*3/5/18*
Targets for today
a) check how stable the rotations advantage is
b) understand and display the mechanism

But I went off that track because I got more to thinking we should be trying to identify circumstances where both strategies 'succeed'. However I suspect that 'success' here is due to low effectiveness and exposure which lead to slow evolution but possible cause low control too.

Looking at trees to differentiate between success, failure and rot_only.
Not very helpful.

Looking at input effects on n generations, they go the way I would expect, but may have to be part of the story.

The violin plots for each input var for rot_seq, rot, neither look pretty good.

e.g. for cost they show that successful runs tend to have higher costs, those failing for both have lower costs. Those that work just for rotations tend to have lower costs.

*8/5/18*

@@

Write out some text explaining the story & maybe the order of the plots. (keep num plots manageable).

order could be
1. example plots
2. histograms 
3. violin plots
4.

Not sure how best to look at effect of inputs on rot_seq, rot, neither. I could bin input ranges and then look at proportions of the 3 within them ?

This would tell me how changing an input value changes the proportion of successes, failures and rot_only. Might be OK to display. Tricky to calculate, but again gets mired in the multivariate nature of the data & potential interactions.

If I go back to using a continuous variable of generations until thresholds I lose out on the successful results.

Maybe a more useful question, rather than when are rotations or sequence better is when are both rotations and sequences sustainable. i.e. under which combination of conditions do both strategies last 'forever' ?

In that case the runs that I have been excluding become the useful ones.

Because actually, finding the runs that are different for rotations and sequences after excluding those that 'succeed' for both is actually trying to find those that 'fail' for sequences.

I think there could be a boundary area in between those that 'succeed' for both and those that 'fail' for both, where some 'succeed' only for rotations.

I think takehome message maybe that rotations are much better than sequences only in rare circumstances.

May also be something useful I can say about the number of insecticides too ? (Which is of interest to IVCC).

Maybe reclass all 10k runs to succeed_both (2), succeed_rot (1), fail_both (0). Then see what are the best predictors of these ? Or do series of xy plots with points coloured, coloured points with all data will probably be a mess ...

~ try changing scenario 2167 from df_cost
~~ initial attempts in UI not working
~~ could it be that it requires just right combination of params and that modifying them just slightly removes the advantage for rotations ?

~ can I create some simple figs to illustrate how an increasing rate of resistance evolution with frequency can lead to rotations being favoured   
~~ maybe just see how dom_cos changes shape of frequency plot
~~ can I look at through UI ?

~ save Ians comments on figY of

~ try decision tree to predict rot_minus_seq from inputs
